---
title: "Lab 6"
author: "Parker Davis"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
    theme: darkly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1

```{r}
getwd()
```
## Task 2

This is the created layout with normal curve plots.:  

```{r}
layout(matrix(1:4, nr=2,nc=2))
#Plot 1
curve(dnorm(x, mean=10,sd=4),xlim=c(-10,30))
#Plot 2
curve(dnorm(x, mean=10,sd=2),xlim=c(0,20))
#Plot 3
curve(dnorm(x, mean=5,sd=10),xlim=c(-30,40))
#Plot 4
curve(dnorm(x, mean=5,sd=.5),xlim=c(3.5,6.5))



```

Individual plots of regions and probabilities:

```{r}
curve(dnorm(x, mean = 0, sd = 1), xlim = c(-3, 3))
xcurve = seq(2, 0+3*1, length = 1000)
ycurve = dnorm(xcurve, mean = 0, sd = 1)
polygon(c(2,xcurve,0+3*1),c(0,ycurve,0),col="Red")
title('Y~ N(0,1), (P >= 2)')
prob1=1-pnorm(2,mean=0,sd=1)
prob1=round(prob1,4)
text(x = 2, y = .35, labels= paste('Prob/Area: ', prob1))
```

```{r}
curve(dnorm(x, mean = 4, sd = 2), xlim = c(-2, 10))
xcurve = seq(1, 5, length = 1000)
ycurve = dnorm(xcurve, mean = 4,sd = 2)
polygon(c(1,xcurve,5),c(0,ycurve,0), col = 'darkgreen')
title('Y ~ N(4,2), P(1 <= Y < 5)')
prob2 = pnorm(5, mean = 4, sd = 2)-pnorm(1, mean = 4, sd = 2)
prob2 = round(prob2, 4)
text(7.6, .17, paste('Prob/Area: ', prob2))
```

```{r}
curve(dnorm(x, mean = 10, sd = 4), xlim = c(-2,22))
xcurve=seq(10-3*4, 10, length = 1000)
ycurve = dnorm(xcurve, mean = 10, sd = 4)
polygon(c(10-3*4,xcurve,10), c(0,ycurve,0), col = 'steelblue')
title('Y ~ N(10,4), P(Y < 10)')
prob3 = pnorm(10, mean = 10, sd = 4)
prob3 = round(prob3, 4)
text(17, 0.08, paste('Prob/Area: ', prob3))
```

```{r}
curve(dnorm(x, mean = -2, sd = .5), xlim = c(-3.5, -.5))
xcurve=seq(-3, -2, length = 1000)
ycurve = dnorm(xcurve, mean = -2, sd = .5)
polygon(c(-3, xcurve, -2), c(0, ycurve, 0), col = 'purple')
title('Y ~ N(-2,.5), P(-3 < Y <= -2)')
prob4 = pnorm(-2, mean = -2, sd=.5)-pnorm(-3, mean = -2, sd = .5)
prob4 = round(prob4, 4)
text(-1, .6, paste('Prob/Area: ', prob4))
```

## Task 3

Overlay of the three gamma plots:  

```{r}
curve(dgamma(x,shape=1,scale=1),xlim=c(0,10),ylim=c(0,1),col="Red",lwd=2,
ylab="Gamma density", main="Gamma curves")
curve(dgamma(x,shape=3,scale=1),xlim=c(0,10),ylim=c(0,1),add=TRUE,lwd=2)
curve(dgamma(x,shape=5,scale=1),xlim=c(0,10),ylim=c(0,1),add=TRUE, col="Blue",lwd=2)
text(6,.8, paste('Red: Shape=1, Black: Shape=3, Blue: Shape=5'))
```

Plotting regions and probabilities:  

```{r}
curve(dgamma(x,shape=3,scale=2),xlim=c(0,20),ylim=c(0,.5),lwd=2,
ylab="Gamma density", main="Y ~ Gamma(shape = 3, scale = 2), P(2 < Y < 5)")
xcurve = seq(2, 5, length = 1000)
ycurve = dgamma(xcurve, shape = 3, scale = 2)
polygon(c(2, xcurve, 5), c(0, ycurve, 0), col = 'red')
gprob1 = pgamma(5, shape=3, scale = 2)-pgamma(2, shape=3, scale = 2)
gprob1 =round(gprob1, 4)
text(15, .4, paste('Prob/Area: ', gprob1))
```

```{r}
curve(dgamma(x,shape=6,scale=3),xlim=c(0,30),ylim=c(0,.09),lwd=2,
ylab="Gamma density", main="Y ~ Gamma(shape = 6, scale = 3), P(1 < Y < 4)")
xcurve = seq(1, 4, length = 1000)
ycurve = dgamma(xcurve, shape = 6, scale = 3)
polygon(c(1, xcurve, 4), c(0, ycurve, 0), col = 'blue')
gprob2 = pgamma(4, shape=6, scale = 3)-pgamma(1, shape=6, scale = 3)
gprob2 =round(gprob2, 4)
text(20, .08, paste('Prob/Area: ', gprob2))
```

```{r}
curve(dgamma(x,shape=2,scale=4),xlim=c(0,30),ylim=c(0,.3),lwd=2,
ylab="Gamma density", main="Y ~ Gamma(shape = 2, scale = 4), P(3 < Y < 6)")
xcurve = seq(3,6, length = 1000)
ycurve = dgamma(xcurve, shape = 2, scale = 4)
polygon(c(3, xcurve, 6), c(0, ycurve, 0), col = 'darkgreen')
gprob3 = pgamma(6, shape=2, scale = 4)-pgamma(3, shape=2, scale = 4)
gprob3 =round(gprob3, 4)
text(20, .08, paste('Prob/Area: ', gprob3))
```

## Task 4

Four chi-square plot layout:

```{r}
layout(matrix(1:4, nr=2,nc=2))

curve(dchisq(x,df=1),xlim=c(0,10),ylim=c(0,1.5),lwd=2,
ylab="Chisq density", main="df=1")
curve(dchisq(x,df=2),xlim=c(0,15),ylim=c(0,1),lwd=2,
ylab="Chisq density", main="df=2")
curve(dchisq(x,df=4),xlim=c(0,20),ylim=c(0,1),lwd=2,
ylab="Chisq density", main="df=4")
curve(dchisq(x,df=20),xlim=c(0,40),ylim=c(0,1),lwd=2,
ylab="Chisq density", main="df=20")
```


Plots with probabilities and regions:  

```{r}
curve(dchisq(x,df=2),xlim=c(0,10),ylim=c(0,1),lwd=2,
ylab="Chisq density", main="Y ~ chisq(df = 2), P(2 <= Y <= 4)")
xcurve = seq(2,4,length = 1000)
ycurve = dchisq(xcurve, df = 2)
polygon(c(2,xcurve,4), c(0,ycurve,0), col = 'pink')
cprob1 = pchisq(4, df =2 ) - pchisq(2, df = 2)
cprob1 = round(cprob1, 4)
text(8,.8, paste('Prob/Area: ', cprob1))
```

```{r}
curve(dchisq(x,df=3),xlim=c(0,15),ylim=c(0,1),lwd=2,
ylab="Chisq density", main="Y ~ chisq(df = 3), P(3 <= Y <= 5)")
xcurve = seq(3,5,length = 1000)
ycurve = dchisq(xcurve, df = 3)
polygon(c(3,xcurve,5), c(0,ycurve,0), col = 'yellow')
cprob2 = pchisq(5, df =3 ) - pchisq(3, df = 3)
cprob2 = round(cprob2, 4)
text(12,.8, paste('Prob/Area: ', cprob2))
```

```{r}
curve(dchisq(x,df=20),xlim=c(0,30),ylim=c(0,1),lwd=2,
ylab="Chisq density", main="Y ~ chisq(df = 20), P(10 <= Y <=21)")
xcurve = seq(10,21,length = 1000)
ycurve = dchisq(xcurve, df = 20)
polygon(c(10,xcurve,21), c(0,ycurve,0), col = 'orange')
cprob3 = pchisq(21, df =20 ) - pchisq(10, df = 20)
cprob3 = round(cprob3, 4)
text(25,.8, paste('Prob/Area: ', cprob3))
```

## Task 5

Four Weibull plots:  

```{r}
layout(matrix(1:4, nr = 2, nc = 2))
curve(dweibull(x,shape=1, scale = 1), xlim=c(0,10), ylab = 'Weibull density', main = 'Shape=1, Scale=1', lwd = 2)

curve(dweibull(x, shape = 4, scale = 1), xlim =c(0,10), ylab = 'Weibull density', main = 'Shape = 4 ,Scale = 1', lwd = 2)

curve(dweibull(x, shape = 3, scale = 2), xlim = c(0,10), ylab = 'Weibull density', main = 'Shape = 3, Scale = 2', lwd = 2)

curve(dweibull(x, shape = 5, scale = 3), xlim = c(0,10), ylab = 'Weibull density', main = 'Shape = 5, Scale = 3', lwd = 2)

```

3 Plots and probabilities:

```{r}
curve(dweibull(x, shape = 3, scale = 5), xlim = c(0,10), ylab = 'Weibull density', main = 'Shape = 3, Scale = 5, P(6<Y<9)', lwd = 2)
xcurve = seq(6,9, length = 1000)
ycurve = dweibull(xcurve, shape = 3, scale = 5)
polygon(c(6,xcurve,9), c(0,ycurve,0), col = 'steelblue')
wprob4 = pweibull(9, shape=3, scale=5)-pweibull(6,shape=3,scale=5)
wprob4=round(wprob4, 4)
text(8,.15,paste('Prob/Area: ', wprob4))
```


```{r}
curve(dweibull(x, shape = 4, scale = 2), xlim = c(0, 10), ylab = 'Weibull density', main = 'Shape = 4, Scale = 2, P(1<Y<2)', lwd = 2)
xcurve = seq(1,2, length = 1000)
ycurve = dweibull(xcurve, shape = 4, scale = 2)
polygon(c(1,xcurve,2), c(0,ycurve,0), col = 'darkgreen')
wprob3 = pweibull(2, shape=4, scale=2)-pweibull(1,shape=4,scale=2)
wprob3=round(wprob3, 4)
text(7,.4,paste('Prob/Area: ', wprob3))
```

```{r}
curve(dweibull(x, shape = 2, scale = 3), xlim = c(0,10), ylab = 'Weibull density', main = 'Shape = 2, Scale = 3, P(2<Y<4)', lwd = 2)
xcurve = seq(2,4, length = 1000)
ycurve = dweibull(xcurve, shape = 2, scale = 3)
polygon(c(2,xcurve,4), c(0,ycurve,0), col = 'purple')
wprob2 = pweibull(4, shape=2, scale=3)-pweibull(2,shape=2,scale=3)
wprob2=round(wprob2, 4)
text(6,.20,paste('Prob/Area: ', wprob2))
```

## Task 6

Four plot layout for the beta function (Limit from 0 to 1):

```{r}
layout(matrix(1:4, nr = 2, nc = 2))
curve(dbeta(x, shape1=1, shape2=1), xlim = c(0,1), ylab = 'Beta density',main = 'Shape1 = 1, Shape2 = 1', lwd =2)
curve(dbeta(x, shape1 = 3, shape2 = 4), xlim = c(0,1), ylab = 'Beta density', main = 'Shape1 = 3, Shape2 = 4', lwd = 2)
curve(dbeta(x, shape1=2, shape2=5), xlim = c(0,1), ylab = 'Beta density',main = 'Shape1 = 2, Shape2 = 5', lwd =2)
curve(dbeta(x, shape1 = 5, shape2 = 3), xlim = c(0,1), ylab = 'Beta density', main = 'Shape1 = 5, Shape2 = 3', lwd = 2)
```

3 plots with probabilities:

```{r}
curve(dbeta(x, shape1 = 1, shape2 = 2), xlim = c(0,1), ylab = 'Weibull density', main = 'Shape1 = 1, Shape2 = 2, P(.2<Y<.4)', lwd = 2)
xcurve= seq(.2, .4, length = 1000)
ycurve = (dbeta(xcurve, shape1 = 1, shape2 = 2))
polygon(c(.2, xcurve,.4), c(0,ycurve,0), col = 'red')
bprob1 = pbeta(.4, shape1 = 1, shape2 = 2)-pbeta(.2, shape1= 1, shape2 = 2)
bprob1 = round(bprob1, 4)
text(.8, 1.2, paste('Prob/Area: ', bprob1))



```

```{r}
curve(dbeta(x, shape1 = 2, shape2 = 4), xlim = c(0,1), ylab = 'Weibull density', main = 'Shape1 = 2, Shape2 = 4, P(.7<Y<.8)', lwd =2)
xcurve= seq(.7, .8, length = 1000)
ycurve = (dbeta(xcurve, shape1 =2, shape2 = 4))
polygon(c(.7, xcurve,.8), c(0,ycurve,0), col = 'steelblue')
bprob2 = pbeta(.8, shape1 = 2, shape2 = 4)-pbeta(.7, shape1= 2, shape2 = 4)
bprop2 = round(bprob2, 4)
text(.8, 1.2, paste('Prob/Area: ', bprob2))
```

```{r}
curve(dbeta(x, shape1 = 3, shape2 = 5), xlim = c(0,1), ylab = 'Weibull density', main = 'Shape1 = 3, Shape2 = 5, P(.3<Y<.5)')
xcurve= seq(.3, .5, length = 1000)
ycurve = (dbeta(xcurve, shape1 = 3, shape2 = 5))
polygon(c(.3, xcurve,.5), c(0,ycurve,0), col = 'pink')
bprob3 = pbeta(.5, shape1 = 3, shape2 = 5)-pbeta(.3, shape1= 3, shape2 = 5)
bprob3 = round(bprob3, 4)
text(.8, 1.2, paste('Prob/Area: ', bprob3))
```

## Task 7

Using the myncurve function:

```{r}
FALL224753davi0607::myncurve(7,10,2)
```

