---
title: "Lab5"
author: "Parker Davis"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1

### Obtain Working Directory
```{r}
getwd()
```

## Task 2

### 'mybin' Function from Dr. Stewart
```{r}

mybin=function(iter=100,n=10, p=0.5){ 
print(p)
sam.mat=matrix(NA,nr=n,nc=iter, byrow=TRUE)

#Create a vector for successes
succ=c()

#loop through all iterations
for( i in 1:iter){

#Fill in the matrix with values of 1 and 0 dependent on p
sam.mat[,i]=sample(c(1,0),n,replace=TRUE, prob=c(p,1-p))

#Sum the successes (1's) from the column i of sam.mat
succ[i]=sum(sam.mat[,i])
}

succ.tab=table(factor(succ,levels=0:n))

barplot(succ.tab/(iter), col=rainbow(n+1), main="Binomial simulation", xlab="Number of successes")
succ.tab/iter

}
```

### 100 Iterations

```{r}
mybin(iter=100,n=10, p=0.7)
```

### 200 Iterations

```{r}
mybin(iter=200,n=10, p=0.7)
```

### 500 Iterations

```{r}
mybin(iter = 500, n = 10, p = 0.7)

```

### 1000 Iterations

```{r}
mybin(iter=1000, n=10, p=0.7)
```

### 10,000 Iterations

```{r}
tab = mybin(iter=10000, n=10, p = 0.7)



```

### 10000 Iteration Table

```{r}
tab
```

### dbinom() Verification

```{r}
```


```{r}
dbin = dbinom(0:10, 10, 0.7)
dbinTab = round((dbin), 4)
names(dbinTab)<- 0:10
dbinTab

```
As we can see, the values given in the table by the dbinom() function closely resemble the values in table we created with the mybin() function!



## Task 3

### Creating a Sample w/o Replacement

```{r}
bagomarbles = function(numMarbles, whiteMarbs, blackMarbs){
marbles <- vector(mode="integer", length = numMarbles)
blackMarbs = blackMarbs + whiteMarbs
for(i in 1:length(marbles))
{
  if (i<=whiteMarbs)
    marbles[i] = 1
  if(i>whiteMarbs & i<=blackMarbs)
    marbles[i] = 0
}
print("1 represents white marbles, 0 represents black marbles")
print("Bag of marbles:")
print(marbles)
}
#bagomarbles(20, 12, 8)
#This is the sample:
sample(bagomarbles(20, 12, 8),size=5, replace=FALSE)

```
### Creating a Sample With Replacement

```{r}
sample(bagomarbles(20,12,8), size = 5, replace = TRUE)
```
### 'myhyper()' Function

```{r}
myhyper=function(iter=100,N=20,r=12,n=5, tab = TRUE){
# make a matrix to hold the samples
#initially filled with NA's
sam.mat=matrix(NA,nr=n,nc=iter, byrow=TRUE)
#Make a vector to hold the number of successes over the trials
succ=c()
for( i in 1:iter){
#Fill each column with a new sample
sam.mat[,i]=sample(rep(c(1,0),c(r,N-r)),n,replace=FALSE)
#Calculate a statistic from the sample (this case it is the sum)
succ[i]=sum(sam.mat[,i])
}
#Make a table of successes
succ.tab=table(factor(succ,levels=0:n))
#Make a barplot of the proportions
barplot(succ.tab/(iter), col=rainbow(n+1), main="HYPERGEOMETRIC simulation", xlab="Number of successes")

#Make an option to display the table of successes (fraction of 1)
if(tab == TRUE)
{
  return(succ.tab/iter)
}

}
```

### Barplots of White Marble Successes over Iteration

```{r}
#Iterate 100 times, total num of marbles = 20, number of whites = 12, take 5 samples
myhyper(iter = 100, N = 20, r = 12, n = 5, tab = FALSE)
myhyper(iter = 200, N = 20, r = 12, n = 5, tab = FALSE)
myhyper(iter = 500, N = 20, r = 12, n = 5, tab = FALSE)
myhyper(iter = 1000, N = 20, r = 12, n = 5, tab = FALSE)
myhyper(iter = 10000, N = 20, r = 12, n = 5, tab = TRUE)
```
### Verifying output

```{r}
dhyper(0:5, 12, 8, 5, log = FALSE)
```
As we can see, the values output by the dhyper() function closely resemble those values that we received from the table of 'succ.tab/iter' from the myhyper() function!

## Task 4

### 'mysample()' Function

Just looking at this function, I can see that it involves taking a random sample of n out of 10 numbers, and then make a factor of that sample. I think that the time part might represent taking a different sample after a certain amount of time, but I am not sure. Maybe it makes as many samples as can fit within the time frame you specify?

```{r}
mysample=function(n, iter=10,time=0.5){
for( i in 1:iter){
#make a sample
s=sample(1:10,n,replace=TRUE)
# turn the sample into a factor
sf=factor(s,levels=1:10)
#make a barplot
barplot(table(sf)/n,beside=TRUE,col=rainbow(10), 
main=paste("Example sample()", " iteration ", i, " n= ", n,sep="") ,
ylim=c(0,0.2)
)

#release the table
Sys.sleep(time)
}
}

```

```{r}
mysample(n = 1000, iter = 1, time = 1)

```


It looks like this mysample() function iterated and created a new sample every 1 to 2 seconds, over the course of 30 iterations. It then created a plot for each iteration, just like with the other functions we have created. It also took 1000 samples of the 1-10 values, and returned in the plot the percentage of successes out of all the responses.  
I would compare the mysample function to dice rolling, as each value has an equal probability. 

NOTE: I could not figure out how to display just one plot out of the 30, so I changed the iterations to just 1 to display only one plot.


## Task 5

### Combination
```{r}
choose(8,4)
```


```{r}
1-ppois(4, lambda = 2)
```

```{r}
dnbinom(7,3,0.4)
```
```{r}
pbinom(8, 15, 0.4)
```


## Task 6

### Bag of Marbles Function
```{r}
FALL224753davi0607::bagomarbles(25,14,11)
newbag <- FALL224753davi0607::bagomarbles(25,14,11)
newbag[1]
```

### mybinom Function

```{r}
FALL224753davi0607::binom(10000,10,.6)
```

```{r}
dbinom(6, 10, .5)
mybin(10000, 10, .5)

((pbinom(6,10,.5)) - (pbinom(3,  10, .5)))

pbinom(6, 10, .5, lower.tail = FALSE)

pbinom(7, 10, .5)


```

