---
title: "Lab 2"
author: "Parker Davis"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1

Working Directory:
```{r}
getwd()
```

## Task 2
```{r}
#Using Dr. Stewart's way
directory="C:\\RPACKAGES\\FALL224753davi0607\\MATH 4753 Labs\\Lab2\\"


#The function to read data in by Dr. Stewart
myread=function(csv){
  fl=paste(directory,csv,sep="")
  read.table(fl,header=TRUE,sep=",")
}


#Assigning maybe in a simpler way to a variable
EPAGAS <- read.csv("EPAGAS.csv")
EPAGASalso <- myread("EPAGAS.csv")
head(EPAGAS, 6)
head(EPAGASalso, 6)

#Ok, so I have two of the same thing but I see how to do it both ways :).
```


## Task 3

```{r}
mpg = EPAGAS$MPG
z=(mpg-mean(mpg))/sd(mpg)
class(z)
#Extremely small mean- approximately 0.
mean(z)
var(z)

#Possible Outliers (Zscores first then Actual Values)
z[abs(z)>=2&abs(z)<=3]
mpg[abs(z)>=2&abs(z)<=3]

#Defined Outliers (Zscores first then Actual Values)
z[abs(z)>3]
mpg[abs(z)>3]

#Creating the column color specifications
labcol= ifelse(abs(z) > 3, "Red", ifelse(abs(z)>=2 & abs(z)<=3, "Blue", "Black"))

#Dotplot of mpg data with OUTLIERS colored.
library(lattice)
dotplot(mpg, col = labcol, main = "MPG Dotplot")
```

## Task 4 

```{r}
#Boxplot for mpg
boxplot(mpg, col = "Black", notch = TRUE, main= "Boxplot representing MPG data",horizontal = TRUE)

```

Using Chebyshev's Theorem, we woud predict the proportion of data within 2 sd of the mean to be (at least):
```{r}
1-(1/(2^2))
```
The exact proportion within 2 sd of the mean is
```{r}
#This takes the 'length'/number of entries within 2 sd of the mean and divides by the actual length to create a proportion
length(mpg[abs(z)<2])/length(mpg)
```

From this information, we can say that Chebyshev would agree with the data. This is because the theorem states that 'at least' 3/4 of the data will fall within 2 sd. While the actual proportion was closer to .96, that still is greater than 3/4, which Chebyshev predicts. Chebyshev's proportion lays out the minimum proportion per amount of standard deviations, so if we had a proportion less than 3/4, then I would be concerned with the data.  
<br>Using the Empirical Rule, approximately .95 is the proportion of the data that should be within 2 standard deviations. This corresponds very well, as the actual proportion of the data is .96.

```{r}
#Verify Empirical Rule Applies
hist(mpg)
```

Thus, I would say that the Empirical Rule is valid, because the two assumptions we make were met:  
1. The dataset is UNIMODAL  
2. The dataset is (approximately) SYMMETRIC  
We can tell this from the boxplot and dotplot that we made. 
