---
title: "Lab 1"
author: "Parker Davis"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1

Working directory is:
```{r}
getwd()
```

## Task 2

First six lines of the DDT.csv is

```{r}
ddt <- read.csv("DDT-1.csv")
head(ddt,6)
```

## Task 3

Qualitative Variables in the DDT data set: River, Species  

Quantitative Variables in the DDT data set: Mile, Length, Weight, DDT  
Number of species in DDT Data set:3 


```{r}
summary.data.frame(ddt) 


big800 = with(ddt, ddt[WEIGHT > 800 & SPECIES == "LMBASS",])

summary(big800)

rvSCM = with(ddt, ddt[RIVER == "SCM"& DDT > 4.0,])

summary(rvSCM)

```

Using these, we can see that the mean length of the fish in the ddt data frame is: 42.81  
The standard deviation for weight of fish in the ddt data set is: 
```{r}
attach(ddt)
sd(WEIGHT)
```



  
Here is the plot of Length vs. Width



```{r}
with(ddt,plot(x = WEIGHT, y = LENGTH, pch=21, bg = "Red", main = "Length vs Width Plot"))
```


## Task 4

Table and barplot of Rivers:

```{r}
rv=with(ddt,table(RIVER))
rv
barplot(rv, beside = TRUE, col = rainbow(4), main = "Rivers")

```

Table and barplot of Rivers crossed with Species of Fish:

```{r}
rvspc=with(ddt,table(RIVER,SPECIES))
rvspc
barplot(rvspc, main = "Table of Rivers crossed with Species of Fish")
```


## Task 5

Pie Charts of Fish Species and Rivers

```{r}
pie(rv, col = 1:4, main = "Rivers")
```

```{r}
spc=with(ddt,table(SPECIES))
pie(spc, col = rainbow(3), main = "Fish Species")
```

## Task 6

These are the boxplots for several variables 

```{r}
with(ddt,boxplot(DDT,ylab = "DDT", col = "Blue", notch = TRUE, main= "DDT Boxplot (with major outlier)"))

with(ddt,boxplot(WEIGHT,ylab="Weight",col="Red",notch=TRUE,main="Weight Boxplot"))

with(ddt, boxplot(LENGTH,ylab="Length",col="Purple",notch = TRUE,main="Length Boxplot"))

```


## Task 7

These are the coplots for several variables

```{r}
coplot(LENGTH~WEIGHT|RIVER,data=ddt,col="Blue")
coplot(DDT~WEIGHT|SPECIES, data=ddt)
```




